Ornaments as indicators of social changes in northeastern Taiwan before and after the European colonial period

### 2020-07-08 analysis/paper/paper.Rmd
minor edit and knit

Archaeological contexts show that ornaments are especially abundant in burial contexts serving as grave goods [@Chen2007]. This supports the interpretation of ornaments as valuable objects functioning as status indicators. Spatial patterns of ornaments in dwelling contexts show that their distribution was clustered during the pre-European and European periods. These clusters are non-random, and are most highly concentrated during the European period. This may indicate that a degree of social inequality based on the uneven distribution of ornaments was already present before European contact, and then it was reinforced and amplified during the European period. A further indicator of increased social inequality is a burial dated to the 17th century that included 60 gold-foil beads, well above the average of 2-3 pieces in the pre-European period [@Chen2007; @Cheng2008]. Based on the finding of that burial, @Cheng2008 proposed a more hierarchical structure of Kiwulan. However, @Hsieh's research on burial data suggested a more egalitarian society since a few burials with rich goods are elders that indicates accumulated wealth over time instead of inherited status. Despite the data they used for making inferences are all based on burial contexts, the inferences about Kiwulan social structure are opposite. Our results can provide an additional insight by focusing on the residential area that demonstrated an uneven distribution of ornaments during the European presence, indicating an increasing social inequality. Nevertheless, we still need to explore and compare with other archaeological records comprehensively discuss the social structure of Kiwulan.

@@

That said, the evidence from Kiwulan may be consistent with a variety of scenarios of indigenous-colonial relations. The increasing number and concentrated spatial patterns of ornaments may also indicate a practice of cultural resistance against the European intrusion. Resistance to European economic and political demands may be inferred if ornaments were used as a display of social identity and to emphasize the local customs that had existed before European contact [cf. @Rubertone2000]. We know the custom of wearing ornaments in Indigenous societies for attending ceremonial events based on the Chinese historical records in the 19th century []. Also, an ethnographic photograph clearly presents the weaving practice by a local woman, who wore a string of beads as a necklace [@Mackey1895, pp. 206-307]. Displaying ornaments in social events or daily live could be a cultural symbol, and may be reinforced during the foreign presence. This could be a form of resistance, but we still need more specific evidence from Kiwulan to prefer resistance as the primary mechanism behind the distribution of ornaments over colonial influence. Another scenario is that ornaments were treated as heirlooms, such as carnelian beads and gold-foil beads, that passed from one generation to the next, accumulating at Kiwulan over time. This process would result in a natural increase in ornaments over time, unrelated to colonial influences. As noted above, this pattern is not a good fit for the ornament distribution at Kiwulan, and adds little value in explaining the shifts in spatial patterns.

@@

By focusing on the distribution patterns in a settlement site, the Kiwulan ornaments suggest that foreign ornaments can be a proxy to detect indirect colonial influence on local Indigenous populations. Ornaments give insights into the amplification of social inequality stimulated by European colonization. It also shows the agency of Indigenous people to incorporate ornaments into their social system and use them in their lives or social occasions to display or intensify status differences. We are still far from understanding the full variety of colonial impacts on peripheral Indigenous communities. We have introduced here the corporate/network model for understanding the dynamics of social inequality at Kiwulan, and further provenance analysis of imported ceramics and ornaments such as X-ray fluorescence analysis would provide more information to construct a clear picture of complex trade networks during this periods.

>>>>

Archaeological contexts show that ornaments are especially abundant in burial contexts serving as grave goods [@Chen2007]. This supports the interpretation of ornaments as valuable objects functioning as status indicators. Spatial patterns of ornaments in dwelling contexts show that their distribution was clustered during the pre-European and European periods. These clusters are non-random, and are most highly concentrated during the European period. This may indicate that a degree of social inequality based on the uneven distribution of ornaments was already present before European contact, and then it was reinforced and amplified during the European period. A further indicator of increased social inequality is a burial dated to the 17th century that included 60 gold-foil beads, well above the average of 2-3 pieces in the pre-European period [@Chen2007; @Cheng2008]. Based on the finding of that burial, @Cheng2008 proposed a more hierarchical structure of Kiwulan. However, @Hsieh2012's research on burial data suggested a more egalitarian society since a few burials with rich goods are elders that indicates accumulated wealth over time instead of inherited status. Despite the data they used for making inferences are all based on burial contexts, the inferences about Kiwulan social structure are opposite. Our results can provide an additional insight by focusing on the residential area that demonstrated an uneven distribution of ornaments during the European presence, indicating an increasing social inequality. Nevertheless, we still need to explore and compare with other archaeological records to make a better inference to social structure of Kiwulan.

@@

That said, the evidence from Kiwulan may be consistent with a variety of scenarios of indigenous-colonial relations. The increasing number and concentrated spatial patterns of ornaments may also indicate a practice of cultural resistance against the European intrusion. Resistance to European economic and political demands may be inferred if ornaments were used as a display of social identity and to emphasize the local customs that had existed before European contact [cf. @Rubertone2000]. We know the custom of wearing ornaments in Indigenous societies for attending ceremonial events based on the Chinese historical records in the 19th century [@Chen1963; @Ke1993]. Also, an ethnographic photo photographed by @Mackey1895, pp. 306-307 presents a weaving practice by an Indigenous woman at Yilan, who wore beaded necklace and headpiece. This indicates displaying ornaments is part of their culture, and may be also used in their daily live according to the photo and the archaeological context where ornaments commonly found in the residential area. This custom might be reinforced during the foreign presence as a form of resistance, but we still need more specific evidence from Kiwulan to prefer resistance as the primary mechanism behind the distribution of ornaments over colonial influence. Another scenario is that ornaments were treated as heirlooms, such as carnelian beads and gold-foil beads, that passed from one generation to the next, accumulating at Kiwulan over time. This process would result in a natural increase in ornaments over time, unrelated to colonial influences. As noted above, this pattern is not a good fit for the ornament distribution at Kiwulan, and adds little value in explaining the shifts in spatial patterns.

@@

By focusing on the distribution patterns in a settlement site, the Kiwulan ornaments suggest that foreign ornaments can be a proxy to detect indirect colonial influence on local Indigenous populations. Ornaments give insights into the amplification of social inequality stimulated by European colonization. It also shows the agency of Indigenous people to incorporate ornaments into their social system and use them in their life or social occasions to display or intensify status differences. We are still far from understanding the full variety of colonial impacts on peripheral Indigenous communities. We have introduced here the corporate/network model for understanding the dynamics of social inequality at Kiwulan, and further provenance analysis of imported ceramics and ornaments such as X-ray fluorescence analysis would provide more information to construct a clear picture of complex trade networks during this periods.
